---
title: "Peak Bloom Prediction Demo"
author: "Siyue Yang"
date: "02/18/2022"
output:
  html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, error = FALSE, 
                      message = FALSE,
                      fig.align = 'center',
                      out.width = '80%')
```

```{r}
library(tidyverse)
library(rnoaa)     # Weather data
library(imputeTS)  # Imputation on the temperature
library(naniar)    # Visulize missing data 
library(cowplot) 
library(lubridate)
```

```{r}
setwd("~/Documents/Github/peak-bloom-prediction")
source("helper_functions.R")
```



# Tokoyo

```{r}
#cherry_jp <- read.csv("data/japan.csv")
all_jp <- read.csv("data/kyoto.csv")
```

```{r}
# Locations near Kyoto
# cherry_jp %>%
#   select(location, lat, long, alt) %>%
#   unique() %>%
#   filter(alt < 45 & alt > 35) %>%
#   ggplot(aes(lat, long, label = location)) +
#   geom_point() + 
#   geom_text(vjust = 0, nudge_y = 0.5)
# 
# cherry_jp %>%
#   filter(location == "Japan/Kumamoto") %>%
#   ggplot(aes(x = year, y = bloom_doy)) +
#   geom_point() +
#   geom_step(linetype = 'dotted', color = 'gray50') +
#   labs(x = "Year", y = "Peak bloom (days since Jan 1st)")
# 
# cherry_jp %>%
#   filter(location == "Japan/Kyoto") %>%
#   ggplot(aes(x = year, y = bloom_doy)) +
#   geom_point() +
#   geom_step(linetype = 'dotted', color = 'gray50') +
#   labs(x = "Year", y = "Peak bloom (days since Jan 1st)")
```

```{r}
# all_jp <- read.csv("data/kyoto.csv")

# We will not use all information after 1950 for Japan data, since the there is no weather data. 
# all_jp %>%
#   ggplot(aes(x = year, y = bloom_doy)) +
#   geom_point() +
#   scale_x_continuous(breaks = seq(810, 2020, by = 40)) +
#   geom_step(linetype = 'dotted', color = 'gray50') +
#   labs(x = "Year", y = "Peak bloom (days since Jan 1st)") +
#   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

all_jp %>%
  filter(year > 1880) %>%
  ggplot(aes(x = year, y = bloom_doy)) +
  geom_point() +
  scale_x_continuous(breaks = seq(1880, 2020, by = 10)) +
  geom_step(linetype = 'dotted', color = 'gray50') +
  labs(x = "Year", y = "Peak bloom (days since Jan 1st)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

# Get weather variables


## GHCND data

GHCND only contains Japan weather data after 1945. In fact, 1945-1950 most of the data are missing. Therefore, we only use data starting from 1950.

```{r}
# options(noaakey = "glXNoFyvOBlaNQHclcFOkcQTFFFLKSdw")
# ncdc_stations(datasetid='GHCND', stationid='GHCND:JA000047759')
```

### Temperature
- tmax = maximum temperature (tenths of degrees C)
- tmin = minimum termperature (tenths of degrees C)
- tavg = average 

### Precipitation 降雨量
- prcp = precipitation (tenths of mm)

### Winter 下雪
- snwd = snow depth, most of them are NA. 


```{r}
cherry.weathers <- 
  tibble(location = "washingtondc", get_seasons.temperature("USC00186350")) %>%
  bind_rows(tibble(location = "liestal", get_seasons.temperature("GME00127786"))) %>%
  bind_rows(tibble(location = "kyoto", get_seasons.temperature("JA000047759"))) %>%
  bind_rows(tibble(location = "vancouver", get_seasons.temperature("CA001108395")))
```

# Data sanity check

1965, 2005 and 2022 data are missing, only contain 1 months data. So we remove the data. 

```{r}
# weather %>%
#   group_by(year) %>%
#   summarise(n = n()) %>%
#   right_join(data.frame(year = c(1950:2022), n = rep(0, 73)), by = "year") %>%
#   filter(is.na(n.x))

# Use 1980s later
'%ni%' <- Negate("%in%")
weather <- weather %>% filter(year >= 1980) %>% filter(year %ni% c(2005, 2022))
```

A lot of them are missing. 

```{r}
vis_miss(weather)
```

## Imputation

```{r}
imp_temperature <- function(dat = weather, temps = "tmin") { 
  
  for (temp in temps) {
    # Convert to time series
    temp_ori <- dat[, temp]
    temp_series <- ts(temp_ori, frequency = 1)
    dat$na <- is.na(temp_ori)
    
    # Use Kalman smoother to impute. Other options are also available
    dat$imp <- na_kalman(temp_series)
    
    n_col <- ncol(dat)
    colnames(dat)[n_col-1] <- paste0(temp, "_na")
    colnames(dat)[n_col] <- paste0(temp, "_imp")
  }
  return(dat)
}
```

```{r}
weather <- imp_temperature(weather, c("tmax", "tavg", "tmin"))
```

A demo of the imputation results. 



```{r}
missing_idx <- which(is.na(weather$tmin))[1:5]
weather$date[missing_idx]

# weather %>%
#   filter(year >= "1964-12-01" & year <= "1972-07-31") %>%
#   ggplot(aes(date, tmin_tmp, color = factor(tmin_na))) +
#   geom_point()

weather %>%
  filter(year >= "1994-12-01" & year <= "1995-07-31") %>%
  ggplot(aes(date, tmin_imp, color = factor(tmin_na))) +
  geom_point()
```

### Temperature

```{r}
get_colour <- function(df){
  colfunc <- colorRampPalette(c("blue", "red"))
  my_colour <- colfunc(12)
   
  df %>%
    group_by(month_name) %>%
    summarise(month_mean = mean(tmax_avg)) %>%
    arrange(month_mean) %>%
    pull(month_name) %>%
    as.integer() -> my_order
   
  my_colour[match(1:12, my_order)]
}
```

```{r}
# https://davetang.org/muse/2020/01/08/plotting-weather-data-using-r/
month_tmax <- weather %>% 
  group_by(month_name, year) %>%
  summarize(tmax_avg = mean(tmax_imp),
            tmin_avg = mean(tmin_imp),
            tavg_avg = mean(tavg_imp),
            prcp_sum = sum(prcp, na.rm = TRUE),
            prcp_avg = mean(prcp, na.rm = TRUE)) 

my_colour <- get_colour(month_tmax %>% select(- tmin_avg, - tavg_avg))
```


```{r}
a <- ggplot(month_tmax, aes(year, tmax_avg, colour = month_name)) +
  geom_point(size = 0.5) +
  geom_smooth(method = "loess") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
        axis.title.x = element_blank(),
        legend.position = "none") +
  scale_color_manual(values = my_colour) +
  labs(title = "Monthly mean maximum temperature", 
       subtitle = "January 1950 - Feburary 2022", y = "Degrees Celsius") +
  facet_wrap(~month_name) +
  NULL

b <- ggplot(month_tmax, aes(year, tmin_avg, colour = month_name)) +
  geom_point(size = 0.5) +
  geom_smooth(method = "loess") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
        axis.title.x = element_blank(),
        legend.position = "none") +
  scale_color_manual(values = my_colour) +
  labs(title = "Monthly mean minimum temperature", 
       subtitle = "January 1950 - Feburary 2022", y = "Degrees Celsius") +
  facet_wrap(~month_name) +
  NULL

plot_grid(a, b)
```

```{r}
a <- month_tmax %>%
  group_by(year) %>%
  summarise(year_mean = mean(tmax_avg)) %>%
  ggplot(., aes(year, year_mean)) +
    geom_point() +
    geom_line() +
    geom_smooth(method = "loess") +
    theme(axis.title.x = element_blank(),
          legend.position = "none") +
    labs(title = "Annual mean maximum temperature", subtitle = ": 1980-2022", y = "Degrees Celsius") +
    NULL
 
b <- month_tmax %>%
  group_by(year) %>%
  summarise(year_mean = mean(tmin_avg)) %>%
  ggplot(., aes(year, year_mean)) +
    geom_point() +
    geom_line() +
    geom_smooth(method = "loess") +
    theme(axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          legend.position = "none") +
    labs(title = "Annual mean minimum temperature", subtitle = ": 1980-2022", y = "Degrees Celsius")  +
    NULL

plot_grid(a, b, nrow = 1, rel_widths = c(1, 1))
```


```{r}
# Scatter plot of low vs high daily temperatures, with a smoother curve for each season
# ggplot(weather, aes(x = tmin_imp, y = tmax_imp)) +
#   geom_point(colour = "firebrick", alpha = 0.1) + 
#   geom_smooth(aes(colour = season), se = F, size = 1.5) +
#   ggtitle ("Daily low and high temperatures") +
#   xlab("Daily low temperature ( ºC )") +  ylab ("Daily high temperature ( ºC )") 
```


```{r}
# https://www.r-bloggers.com/2015/03/part-3a-plotting-with-ggplot2/
# Distribution of the average temperature by season - density plot
season_tmax <- weather %>% 
  group_by(season, year) %>%
  summarize(tmax_avg = mean(tmax, na.rm= TRUE),
            tmin_avg = mean(tmin, na.rm= TRUE),
            tavg_avg = mean(tavg, na.rm= TRUE)) 

ggplot(season_tmax, aes(x = tmax_avg, colour = season)) +
  geom_density() +
  scale_x_continuous(limits = c(5,30), breaks = seq(5,30,5)) +
  ggtitle ("Temperature distribution by season") +
  xlab("Average temperature ( ºC )") +  ylab ("Probability") 
```

#### Freezing dates

"Gardeners will keep an eye on the weather forecast, and keen gardeners will want to understand historical trends as well. An area of particular interest is the last freeze each spring. If seeds or starts are planted too early, they may succumb to a late frost. On the other end of the growing season, the first freeze each fall is a good indication of when to ensure the harvest is completed for crops that cannot withstand an early frost."

```{r}
# Accumulative freezing days
# Return a dataframe with year as the index variable and 
get_afdd <- function(dat) {
  
  freeze_days <- dat %>% mutate(yday = yday(date)) %>% group_by(year) %>% filter(tmin_imp <= 0)
  
  # Number of accumulative freezing days
  freeze_days_n <- freeze_days %>% group_by(year) %>% summarise(afdd = n()) 
  
  # First and last freeze
  first_freeze <- freeze_days %>% group_by(year) %>% summarise(first_freeze_date = min(date))
  last_freeze <- freeze_days %>% group_by(year) %>% summarise(last_freeze_date = max(date))

  first_freeze_yday <- first_freeze %>% mutate(first_freeze_yday = yday(first_freeze_date))
  last_freeze_yday <- last_freeze %>% mutate(last_freeze_yday = yday(last_freeze_date))      
  
  year_range <- c(min(dat$year) : max(dat$year))
  
  freeze_info <- merge(data.frame(year = year_range), freeze_days_n, all.x = TRUE) 
  freeze_info <- merge(freeze_info, first_freeze_yday, by = "year", all.x = TRUE)
  freeze_info <- merge(freeze_info, last_freeze_yday, by = "year", all.x = TRUE)
  
  return(freeze_info)
}

get_freeze_days <- function(dat) {
  freeze_days <- dat %>% mutate(yday = yday(date)) %>% group_by(year) %>% filter(tmin_imp <= 0)
  return(freeze_days)
}
```



```{r}
# Accumulative growing days
# Average temperature = (tmax + tmin)/2 > 0
get_agdd <- function(dat) {
  
  growing_days <- dat %>% mutate(yday = yday(date)) %>% group_by(year) %>% 
    mutate(tavg_cal = (tmax_imp + tmin_imp)/2) %>% filter(tavg_cal > 0)
  
  # Number of accumulative growing days
  grow_days_n <- growing_days %>% group_by(year) %>% summarise(agdd = n()) 
  
  # First and last freeze
  first_grow <- growing_days %>% group_by(year) %>% summarise(first_grow_date = min(date))
  last_grow <- growing_days %>% group_by(year) %>% summarise(last_grow_date = max(date))

  first_grow_yday <- first_grow %>% mutate(first_grow_yday = yday(first_grow_date))
  last_grow_yday <- last_grow %>% mutate(last_grow_yday = yday(last_grow_date))      
  
  year_range <- c(min(dat$year) : max(dat$year))
  
  freeze_info <- merge(data.frame(year = year_range), grow_days_n, all.x = TRUE) 
  freeze_info <- merge(freeze_info, first_grow_yday, by = "year", all.x = TRUE)
  freeze_info <- merge(freeze_info, last_grow_yday, by = "year", all.x = TRUE)
  
  return(freeze_info)
}
```

```{r}
get_agdd(weather)
```


```{r}

# # Get all freeze days, tmin < 0 C
# freeze_days <- weather %>% mutate(yday = yday(date)) %>% group_by(year) %>% filter(tmin_imp <= 0) 
# 
# # First and last freeze
# first_freeze <- freeze_days %>% group_by(year) %>% summarise(min = min(date))
# last_freeze <- freeze_days %>% group_by(year) %>% summarise(max = max(date))
# 
# first_freeze_yday <- first_freeze %>% mutate(yday = yday(min))
# last_freeze_yday <- last_freeze %>% mutate(yday = yday(max))           


adgg_jp <- get_afdd(weather)
freeze_days <- get_freeze_days(weather)

ggplot() +
  geom_point(data = freeze_days, aes(y = year, x = as.Date(yday, origin='2022-01-01')), 
             color = "gray", size = 0.8) +
  geom_point(data = adgg_jp, aes(y=year, x = as.Date(first_freeze_yday, origin='2022-01-01')),
             fill = "purple",shape = 21, size = 3, color = 'black') +
  geom_point(data = adgg_jp, aes(y = year, x = as.Date(last_freeze_yday, origin='2022-01-01')) 
             ,fill = "blue",shape = 21, size = 3, color = 'black') +
  scale_y_reverse(breaks = seq(1950, 2022, 10), minor_breaks = seq(1944, 2022, 2)) +
  scale_x_date(date_labels = "%b", breaks = "month") +
  ggtitle("Yearly Last Freeze (Spring) and First Freeze (Winter) Dates") +
  xlab("Date")
```
Blue dots indicate the last freeze of winter
Purple dots indicate the first freeze of fall
Gray dots indicate days with a min temp at or below 0 tenths of degrees C


```{r}
# 1965 - 1972 no freezing days
freeze_days %>%
  group_by(year) %>%
  summarise(n = n()) %>%
  right_join(data.frame(year = c(1980:2021), n = rep(0, 42)), by = "year") %>%
  mutate(n = n.x + n.y) %>%
  left_join(all_jp, by = "year") %>%
  ggplot() +
  geom_bar(aes(x = year, y = n), stat = "identity") +
  geom_point(aes(x = year, y = bloom_doy))
  xlab("Number of freezing days")
```

- 最近10年最后化冰日提前了，大概在2月底。
- 有的时候结冰期（紫色）退后了。

### Rainfall

```{r}
weather %>%
  group_by(year, month) %>%
  summarise(prcp_sum = sum(prcp, na.rm = T)) %>%
  ggplot(aes(x = factor(month), y = prcp_sum)) +
  geom_boxplot() +
  theme_classic() +
  ggtitle("Cumulative rainfall by month at Kyoto (1953 - 2022)") 

weather %>%
  group_by(year, month) %>%
  summarise(prcp_sum = mean(prcp, na.rm = T)) %>%
  ggplot(aes(x = year, y = prcp_sum)) +
  geom_point() +
  theme_classic() +
  ggtitle("Yearly inches of rainfall by month at Kyoto (1953 - 2022)") 
```

```{r}
# weather %>% 
#   mutate(prcp_avg_all = mean(prcp, na.rm = T)) %>%
#   mutate(prcp_pos = case_when(prcp - prcp_avg_all > 0 ~ prcp - prcp_avg_all, TRUE ~ 0),
#          prcp_neg = case_when(prcp - prcp_avg_all < 0 ~ prcp - prcp_avg_all, TRUE ~ 0)) %>%
#   ggplot(aes(x = year)) + 
#   geom_ribbon(aes(ymin = 0, ymax = prcp_pos, fill="positive")) +
#   geom_ribbon(aes(ymin = prcp_neg, ymax = 0, fill="negative")) +
#   scale_fill_manual(name="",values=c("#D6604D", "#4393C3")) +
#   ylim(-100, 100) +
#   ylab("rainfall anomaly in mm")
```


# Modeling

```{r}
cherry <- read.csv("data/washingtondc.csv") %>% 
  bind_rows(read.csv("data/liestal.csv")) %>% 
  bind_rows(read.csv("data/kyoto.csv"))

model_data <- cherry.weathers %>% merge(cherry, by = c("year", "location"), all.y = TRUE, all.x = TRUE) %>%
  filter(year >= 1950)
```

```{r}
model_data_jp <- model_data %>% 
  filter(location == "kyoto") %>%
  drop_na()
```



```{r}
library(corrplot)
corr_data <- model_data_jp %>% select(where(is.numeric)) %>%
  select(-c(lat, long, alt)) 

corr_data <- cor(corr_data)
corrplot(corr_data)
```


```{r}
# mod1 <- lm(bloom_doy ~ year + afdd + agdd + 
#              tmax_avg_Winter + tmax_avg_Spring + tmin_avg_Spring + prcp_avg_Winter + 
#                first_grow_yday + first_freeze_yday , data = model_data_jp)
# summary(mod1)


# https://www.r-bloggers.com/2017/01/doing-magic-and-analyzing-seasonal-time-series-with-gam-generalized-additive-model-in-r/
library(mgcv)
# mod2 <- gam(bloom_doy ~ s(tmax_avg_Spring, bs = "ps", k = 7) + tmin_avg_Spring, data = model_data_jp)
# summary(mod2)

gam.m3 <- mgcv::gam(bloom_doy ~ s(afdd)  +
             s(tmax_avg_Winter) + s(tmax_avg_Spring) + s(prcp_avg_Winter) + s(tmin_avg_Spring) +
                s(prcp_avg_Spring),
              data = model_data_jp)
layout(matrix(1:6, nrow = 3))
plot(gam.m3, shade = TRUE)

summary(gam.m3)

```

```{r}
library(data.table)

# datas <- rbindlist(list(model_data_jp[, c("bloom_doy", "year")],
#                         data.table(value = mod2$fitted.values,
#                                    data_time = model_data_jp[, "year"])))
# datas[, type := c(rep("Real", nrow(model_data_jp)), rep("Fitted", nrow(model_data_jp)))]
#  
# ggplot(data = datas, aes(year, bloom_doy, group = type, colour = type)) +
#   geom_line(size = 0.8) +
#   theme_bw() +
#   labs(x = "Time", y = "Day of year",
#        title = "Fit from GAM n.1")

datas <- rbindlist(list(model_data_jp[, c("bloom_doy", "year")],
                        data.table(value = gam.m3$fitted.values,
                                   data_time = model_data_jp[, "year"])))
datas[, type := c(rep("Real", nrow(model_data_jp)), rep("Fitted", nrow(model_data_jp)))]
 
ggplot(data = datas, aes(year, bloom_doy, group = type, colour = type)) +
  geom_line(size = 0.8) +
  theme_bw() +
  labs(x = "Time", y = "Day of year",
       title = "Fit from GAM n.1")
```

```{r}
gam.m3 <- mgcv::gam(bloom_doy ~ s(tmax_avg_Winter) + s(tmax_avg_Spring) +  
                      t2(tmin_avg_Spring) + s(prcp_avg_Spring),
              data = model_data_jp)

gam1 <- mgcv::gam(bloom_doy ~ s(afdd, prcp_avg_Spring) + s(tmax_avg_Winter) + s(tmin_avg_Spring, tmax_avg_Spring), data = model_data_jp)

# year + afdd + agdd + 
#              tmax_avg_Winter + tmax_avg_Spring + tmin_avg_Spring + prcp_avg_Winter + 
#                first_grow_yday + first_freeze_yday , data = model_data_jp)
# layout(matrix(1:2, nrow = 4))
# plot(gam.m3, shade = TRUE)

summary(gam.m3)$r.sq
summary(gam1)$r.sq

# AIC(gam.m3, gam1)
```

```{r}
ggplot(model_data_jp) +
  geom_point(aes(afdd, bloom_doy, color = tmax_avg_Winter)) 
```



```{r}
gam.check(gam.m3)
```


# Prediction

```{r}
train_data <- model_data_jp[1:60, ]
test_data <- model_data_jp[61:70, ]
```

```{r}
train_mod <- mgcv::gam(bloom_doy ~ s(afdd, prcp_avg_Spring) + s(tmax_avg_Winter) + s(tmin_avg_Spring, tmax_avg_Spring), data = model_data_jp)

summary(train_mod)

y_hat_train <- predict(train_mod, data = train_data)
train_MSE <- mean(abs(train_data$bloom_doy - round(y_hat_train)))

y_hat_test <- predict(train_mod, newdata = test_data)
test_MSE <- mean(abs(test_data$bloom_doy - round(y_hat_test)))

nonp <- round(y_hat_test)
train_MSE
test_MSE
# ?cross-validation?
```

```{r}
library(randomForest)
# random forest
rf <- randomForest(bloom_doy ~  afdd +  tmax_avg_Winter + tmax_avg_Spring + tmin_avg_Spring + prcp_avg_Winter, data=train_data, mtry = 4/3, importance = TRUE, ntrees = 500)
rf

varImpPlot(rf, main = "Variable importance")
```

```{r}

# pred_rf <- predict(rf_model, data.frame(matrix_test)) + mean(trend_for)
# 
# 
# pred_rf <- data.table(value = pred_rf, Var2 = 1:48, Var1 = "RF")
#  
# pred_true <- data.table(value = data_test$value, Var2 = 1:48, Var1 = "Real")
# preds_all <- rbindlist(list(pred_ave_rpart, pred_ave_ctree, pred_rf, pred_true), use.names = T)
#  
# ggplot(preds_all, aes(Var2, value, color = as.factor(Var1))) +
#   geom_line(alpha = 0.7, size = 1.2) +
#   labs(x = "Time", y = "Load (kW)", title = "Comparison of Ensemble Learning forecasts") +
#   guides(color=guide_legend(title="Method")) +
#   theme_ts
```


```{r}
rf <- round(predict(rf, test_data))
shiki <- c(97,90,88,85,82,88,84,87,86,87)
avg <- round((rf + shiki + nonp)/3)
cbind(true = test_data$bloom_doy, rf, shiki, nonp, avg)

cal_abe <- function(true, test) {
  mean(abs(true - test))
}


cal_abe(test_data$bloom_doy, rf)
cal_abe(test_data$bloom_doy, shiki)
cal_abe(test_data$bloom_doy, nonp)
cal_abe(test_data$bloom_doy, avg)
```



```{r}
model_data_li <- model_data %>% 
  filter(location == "liestal") %>%
  drop_na()
```

```{r}
train_data <- model_data_li[1:56, ]
test_data <- model_data_li[57:66, ]

train_mod <- mgcv::gam(bloom_doy ~ s(prcp_avg_Spring) + s(tmax_avg_Winter) + s(tmin_avg_Spring, tmax_avg_Spring), data = model_data_li)

y_hat_train <- predict(train_mod, data = train_data)
train_MSE <- mean(abs(train_data$bloom_doy - round(y_hat_train)))

y_hat_test <- predict(train_mod, newdata = test_data)
test_MSE <- mean(abs(test_data$bloom_doy - round(y_hat_test)))

train_MSE
test_MSE
```

```{r}
library(randomForest)
# random forest
rf <- randomForest(bloom_doy ~  afdd +  tmax_avg_Winter + tmax_avg_Spring + tmin_avg_Spring + prcp_avg_Winter, data=train_data, mtry = 4/3, importance = TRUE, ntrees = 500)
rf

```


```{r}
nonp <- round(y_hat_test)
rf <- round(predict(rf, test_data))
shiki <- c(96, 70, 77, 107, 72, 87, 85, 86, 71, 84)
avg <- round((rf + shiki + nonp)/3)
cbind(true = test_data$bloom_doy, rf, shiki, nonp, avg)

cal_abe <- function(true, test) {
  mean(abs(true - test))
}


cal_abe(test_data$bloom_doy, rf)
cal_abe(test_data$bloom_doy, shiki)
cal_abe(test_data$bloom_doy, nonp)
cal_abe(test_data$bloom_doy, avg)
```


```{r}
model_data_dc <- model_data %>% 
  filter(location == "washingtondc") %>%
  drop_na()
```

```{r}
train_data <- model_data_dc[1:60, ]
test_data <- model_data_dc[61:70, ]

train_mod <- mgcv::gam(bloom_doy ~ s(prcp_avg_Spring) + s(tmax_avg_Winter) + s(tmin_avg_Spring, tmax_avg_Spring), data = model_data_dc)

y_hat_train <- predict(train_mod, data = train_data)
train_MSE <- mean(abs(train_data$bloom_doy - round(y_hat_train)))

y_hat_test <- predict(train_mod, newdata = test_data)
test_MSE <- mean(abs(test_data$bloom_doy - round(y_hat_test)))

train_MSE
test_MSE
```

```{r}
library(randomForest)
# random forest
rf <- randomForest(bloom_doy ~  afdd +  tmax_avg_Winter + tmax_avg_Spring + tmin_avg_Spring + prcp_avg_Winter, data=train_data, mtry = 4/3, importance = TRUE, ntrees = 500)
rf

```

```{r}
nonp <- round(y_hat_test)
rf <- round(predict(rf, test_data))
shiki <- c(85, 88, 90, 88, 83, 85, 92, 90, 80, 89)
avg <- round((rf + shiki + nonp)/3)
cbind(true = test_data$bloom_doy, rf, shiki, nonp, avg)

cal_abe <- function(true, test) {
  mean(abs(true - test))
}


cal_abe(test_data$bloom_doy, rf)
cal_abe(test_data$bloom_doy, shiki)
cal_abe(test_data$bloom_doy, nonp)
cal_abe(test_data$bloom_doy, avg)
```

SVM

```{r}
library(e1071)
svmfit = svm(bloom_doy ~ afdd + tmax_avg_Winter + tmax_avg_Spring + tmin_avg_Spring + prcp_avg_Winter, data=train_data)
summary(svmfit)

svm_predict <- round(predict(svmfit, test_data))

cal_abe(train_data$bloom_doy, svm_predict)
cal_abe(test_data$bloom_doy, svm_predict)

```



# Forecast


```{r}
dailyweather <- get_ghcnd(station_id = "JA000047759", info_list = c("tmax", "tmin", "prcp"), date_range = c("1950-01-01", "2022-01-31"))

dailyweather <- imp_temperature(dailyweather, c("tmax", "tmin"))

forecast_jp <- forecast_weather(dailyweather)
```




Of course, the data has seasonality.... 

1. benchmark methods to forecast

$$y_t = y_{t-s} + e_t$$

```{r}
# residual sd = 1.7848 
fit <- snaive(ts_jp)
print(summary(fit))
checkresiduals(fit)
```


2. Fit ETS method, classic time series forcasting model

```{r}
fit_ets <- ets(ts_jp)  # SD =  1.735
print(summary(fit_ets))
checkresiduals(fit_ets)
```

3. Fit ARIMA model

```{r}
fit_arima <- auto.arima(ts_jp)  # SD = 1.421267, Ljung-Box test p-val = 0.57
print(summary(fit_arima))
checkresiduals(fit_arima)
```

```{r}
fcst <- forecast(fit_arima, h = 120)
autoplot(fcst, include = 100)
```




## Cluster

```{r}
library(mclust)

cherry_k <- read.csv("data/south_korea.csv") %>% mutate(Country = "Korea")
cherry_s <- read.csv("data/meteoswiss.csv") %>% mutate(Country = "Swiss")
cherry_j <- read.csv("data/japan.csv") %>% mutate(Country = "Japan")

cherry_all <- cherry_k %>% bind_rows(cherry_s) %>% bind_rows(cherry_j)
  

cherry_all %>% 
  ggplot() +
  geom_point(aes(x = year, y = bloom_doy, color = Country)) +
  xlim(c(1950, 2021)) 

cherry_all %>% 
  ggplot() +
  geom_density(aes(x = bloom_doy, fill = Country)) 

mod1 <- Mclust(cherry_all$bloom_doy)

summary(mod1, parameters = T)
```

```{r}
pheno <- read_csv("data/USA-NPN_status_intensity_observations_data.csv")

pheno %>%
  group_by(Genus, Species, Common_Name ) %>%
  summarise(n = n())


pheno$date <- as.POSIXct(pheno$Observation_Date, format = "%m/%d/%Y")
pheno$year <- format(pheno$date, format = "%Y")

yedoensis <- pheno %>%
  filter(Species == "yedoensis") %>%
  group_by(State, year) %>%
  mutate(n = n()) %>%
  filter(n >= 10) %>%
  unique() %>%
  filter(Day_of_Year >= quantile(Day_of_Year, 0.69) & Day_of_Year <= quantile(Day_of_Year, 0.71))
  

yedoensis

yedoensis %>%
  ggplot() +
  geom_point(aes(x = factor(year), y = as.integer(Day_of_Year), color = factor(State)))

yedoensis %>%
  ggplot() +
  geom_density(aes(x = Day_of_Year))

mod1 <- Mclust(yedoensis$Day_of_Year)

summary(mod1, parameters = T)

```

```{r}
date_van <- c("03/31/2004", "04/09/2005", "03/30/2006", "04/01/2007", "03/26/2008", "04/01/2009", "03/31/2010", "03/29/2011", "03/20/2012", "04/09/2013", "04/10/2014", "04/10/2015", "03/25/2016", "03/25/2017", "04/05/2018", "04/01/2019", "03/20/2020", "03/28/2021")

date_van <- data.frame(date = date_van) %>%
  mutate(date = as.POSIXct(date, format = '%m/%d/%Y')) %>%
  mutate(year = as.integer(format(date, "%Y"))) %>%
  mutate(bloom_doy = yday(date)) %>%
  mutate(location = "vancouver")

ggplot() +
  geom_point(data = model_data, aes(x = year, y = bloom_doy, color = location)) +
  geom_point(data = date_van, aes(x = year, y = bloom_doy, color = "pink"))

cherry_all <- model_data %>%
  select(year, bloom_doy, location) %>%
  bind_rows(date_van %>% select(year, bloom_doy, location)) 
  
cherry_all %>%
  ggplot() +
  geom_boxplot(aes(x = location, y = bloom_doy))

cherry_all %>%
  ggplot() +
  geom_density(aes(x = bloom_doy, color = location))
```

```{r}
cherry <- read.csv("data/japan.csv") %>% 
  mutate(Country = "Jp") %>%
  bind_rows(read.csv("data/meteoswiss.csv") %>% mutate(Country = "Sw")) %>% 
  bind_rows(read.csv("data/south_korea.csv") %>% mutate(Country = "Kr")) 

ggplot(data = cherry) +
  geom_density(aes(x = bloom_doy, color = Country))
```

Semi-supervised learning

```{r}
library(ssr)

train_data <- model_data_vc[1:40, ] 
train_data <- train_data %>% 
  select(bloom_doy, afdd, tmax_avg_Winter, tmax_avg_Spring, prcp_avg_Winter)

test_data <- model_data_vc[41:47, ] 
test_data <- test_data %>%
  select(bloom_doy, afdd, tmax_avg_Winter, tmax_avg_Spring, prcp_avg_Winter)

set.seed(1234)

# # Split the dataset into 70% for training and 30% for testing.
# split1 <- split_train_test(dataset, pctTrain = 70)
# 
# # Choose 5% of the train set as the labeled set L and the remaining will be the unlabeled set U.
# split2 <- split_train_test(split1$trainset, pctTrain = 5)


label_index <- which(is.na(train_data$bloom_doy) == 1)
L <- train_data[-label_index, ] # This is the labeled dataset.

U <- train_data[label_index, -1] # Remove the labels since this is the unlabeled dataset.

testset <- test_data %>% drop_na() # This is the test set.
```

```{r}
regressors <- list(linearRegression=lm, knn=caret::knnreg, svm=e1071::svm)

# Fit the model.
model <- ssr("bloom_doy ~ .", L, U, regressors = regressors, testdata = testset)
```

```{r}
U
```



```{r}
regressors <- list("lm", "rvmLinear")
```

```{r}
plot(model)

# Get the predictions on the testset.
predictions <- predict(model, testset)

# Calculate RMSE on the test set.
rmse.result <- sqrt(mean((predictions - testset$bloom_doy)^2))
rmse.result
#> [1] 0.1113865

plot(model, metric = "mae", ptype = 2)
```

```{r}
train_mod <- mgcv::gam(bloom_doy ~ s(tmax_avg_Spring) , data = L)

y_hat_test <- predict(train_mod, newdata = testset)

nonp <- round(y_hat_test)

cal_abe(testset$bloom_doy, round(predictions))

cal_abe(testset$bloom_doy, nonp)

cbind(round(predictions), nonp, testset$bloom_doy)


```

